---
title: "nof1 disease up outlier thresholds"
author: "Kaylee Delgado"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

## Set Up Library Packages
```{r setup}

knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(readr)
library(tidyr)
library(tidyverse)
library(stringr)
library(conflicted)
library(kableExtra)
library(jsonlite)
library(knitr)

conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::intersect)
```

## Read json file with down outlier thresholds
```{r input json}

# Construct the file path.  This is important for portability!
json_file3.0 <- file.path(getwd(), "3.json")

# Read the JSON data
json_data3.0 <- fromJSON(json_file3.0)
```


## Input Up Outlier Threshold Data
```{r input up outlier threshold data}

# Extract up outlier nof1 genes from 3json file (lets call, data set c)
threshold_up_outlier_genes_json <- json_data3.0$nof1_disease_thresholds$high

# Create tibble with up outlier gene name and threshold value
threshold_up_outlier_genes_df <- tibble(gene=names(threshold_up_outlier_genes_json),
       threshold=unlist(threshold_up_outlier_genes_json))

# Style kable table 
kable(threshold_up_outlier_genes_df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


# Up Outliers Samples 1-6: Gene, Expression, and Threshold

## Create Up Outlier Gene tibble for Sample 1, including gene and expression level (4.0json), and merge with tibble of extracted outliers (nof1disease) from 4.0json
```{r input up gene data sample 1}

# 1. Extract only up outlier data from nof1 outlier results for sample 1 (lets call, data set A) (from "Sample-1-to-6-nof1-outlier.rmd)
nof1_up_data_1 <- read_tsv("Sample 1-6 nof1 Outlier Results.tsv", show_col_types = FALSE) %>%
  filter(direction == "up") %>%
  filter(TH_id == "TH50_5178_S01") %>% 
  filter(!grepl("^RP", gene)) # filter out genes starting with "RP"
  
# Add expression levels from outlier_results$sample (lets call, data set B)
# 2. Extract expression data from a JSON file 4.0
json_file4.0_1 <- file.path(getwd(), "TH50_5178_S01.4.0.json")

json_data4.0_1 <- fromJSON(json_file4.0_1)

up_outlier_expression_df_1 <- tibble(gene=names(json_data4.0_1$outlier_results$sample), expression=unlist(json_data4.0_1$outlier_results$sample))

print(up_outlier_expression_df_1)

# 3. Merge the two datasets (inner_join for overlapping genes) (A+B)
up_merged_data_1 <- inner_join(nof1_up_data_1, up_outlier_expression_df_1, by = "gene")

# 4. Select columns 
up_final_data_1 <- up_merged_data_1 %>%
  select(gene, direction, expression)

print(up_final_data_1)

# 5. Merge the threshold data with the expression data
up_final_merged_data_1 <- inner_join(up_final_data_1, threshold_up_outlier_genes_df, by = "gene") %>%
  select(gene, expression, threshold, direction)

# Style kable table 
kable(up_final_merged_data_1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "500px")

# Export as TSV
write.table(up_final_merged_data_1, "Sample-1-Up-Outliers-RP-Filtered.tsv", sep = "\t", row.names = FALSE, quote = FALSE)
```

## Create Up Outlier Gene tibble for Sample 2, including gene and expression level (4.0json), and merge with tibble of extracted outliers (nof1disease) from 4.0json
```{r input up gene data sample 2}

# 1. Repeat step #1 for sample 2, changing TH_id
nof1_up_data_2 <- read_tsv("Sample 1-6 nof1 Outlier Results.tsv", show_col_types = FALSE) %>%
  filter(direction == "up") %>%
  filter(TH_id == "TH50_5179_S01") %>% 
  filter(!grepl("^RP", gene))
  
# Add expression levels from outlier_results$sample (lets call, data set B)
# 2. Repeat step 2, changing TH_id
json_file4.0_2 <- file.path(getwd(), "TH50_5179_S01.4.0.json")

json_data4.0_2 <- fromJSON(json_file4.0_2)

up_outlier_expression_df_2 <- tibble(gene=names(json_data4.0_2$outlier_results$sample), expression=unlist(json_data4.0_2$outlier_results$sample))

print(up_outlier_expression_df_2)

# 3. Merge the two datasets (inner_join for overlapping genes) (A+B)
up_merged_data_2 <- inner_join(nof1_up_data_2, up_outlier_expression_df_2, by = "gene")

# 4. Select columns 
up_final_data_2 <- up_merged_data_2 %>%
  select(gene, direction, expression)

print(up_final_data_2)

# 5. Merge the threshold data with the expression data
up_final_merged_data_2 <- inner_join(up_final_data_2, threshold_up_outlier_genes_df, by = "gene") %>%
  select(gene, expression, threshold, direction)

# Style kable table 
kable(up_final_merged_data_2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "500px")

# Export as TSV
write.table(up_final_merged_data_2, "Sample-2-Up-Outliers-RP-Filtered.tsv", sep = "\t", row.names = FALSE, quote = FALSE)
```

## Create Up Outlier Gene tibble for Sample 3, including gene and expression level (4.0json), and merge with tibble of extracted outliers (nof1disease) from 4.0json
```{r input up gene data sample 3}

# 1. Repeat step #1 for sample 3, changing TH_id
nof1_up_data_3 <- read_tsv("Sample 1-6 nof1 Outlier Results.tsv", show_col_types = FALSE) %>%
  filter(direction == "up") %>%
  filter(TH_id == "TH50_5280_S01") %>% 
  filter(!grepl("^RP", gene))
  
# Add expression levels from outlier_results$sample (lets call, data set B)
# 2. Repeat step 2, changing TH_id
json_file4.0_3 <- file.path(getwd(), "TH50_5280_S01.4.0.json")

json_data4.0_3 <- fromJSON(json_file4.0_3)

up_outlier_expression_df_3 <- tibble(gene=names(json_data4.0_3$outlier_results$sample), expression=unlist(json_data4.0_3$outlier_results$sample))

print(up_outlier_expression_df_3)

# 3. Merge the two datasets (inner_join for overlapping genes) (A+B)
up_merged_data_3 <- inner_join(nof1_up_data_3, up_outlier_expression_df_3, by = "gene")

# 4. Select columns 
up_final_data_3 <- up_merged_data_3 %>%
  select(gene, direction, expression)

print(up_final_data_3)

# 5. Merge the threshold data with the expression data
up_final_merged_data_3 <- inner_join(up_final_data_3, threshold_up_outlier_genes_df, by = "gene") %>%
  select(gene, expression, threshold, direction)

# Style kable table 
kable(up_final_merged_data_3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "500px")

# Export as TSV
write.table(up_final_merged_data_3, "Sample-3-Up-Outliers-RP-filtered.tsv", sep = "\t", row.names = FALSE, quote = FALSE)

```

## Create Up Outlier Gene tibble for Sample 4, including gene and expression level (4.0json), and merge with tibble of extracted outliers (nof1disease) from 4.0json
```{r input Up gene data sample 4}

# 1. Repeat step #1 for sample 4, changing TH_id
nof1_up_data_4 <- read_tsv("Sample 1-6 nof1 Outlier Results.tsv", show_col_types = FALSE) %>%
  filter(direction == "up") %>%
  filter(TH_id == "TH50_5281_S01") %>% 
  filter(!grepl("^RP", gene))
  
# Add expression levels from outlier_results$sample (lets call, data set B)
# 2. Repeat step 2, changing TH_id
json_file4.0_4 <- file.path(getwd(), "TH50_5281_S01.4.0.json")

json_data4.0_4 <- fromJSON(json_file4.0_4)

up_outlier_expression_df_4 <- tibble(gene=names(json_data4.0_4$outlier_results$sample), expression=unlist(json_data4.0_4$outlier_results$sample))

print(up_outlier_expression_df_4)

# 3. Merge the two datasets (inner_join for overlapping genes) (A+B)
up_merged_data_4 <- inner_join(nof1_up_data_4, up_outlier_expression_df_4, by = "gene")

# 4. Select columns 
up_final_data_4 <- up_merged_data_4 %>%
  select(gene, direction, expression)

print(up_final_data_4)

# 5. Merge the threshold data with the expression data
up_final_merged_data_4 <- inner_join(up_final_data_4, threshold_up_outlier_genes_df, by = "gene") %>%
  select(gene, expression, threshold, direction)

# Style kable table 
kable(up_final_merged_data_4) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "500px")

# Export as TSV
write.table(up_final_merged_data_4, "Sample-4-Up-Outliers-RP-Filtered.tsv", sep = "\t", row.names = FALSE, quote = FALSE)

```

## Create Up Outlier Gene tibble for Sample 5, including gene and expression level (4.0json), and merge with tibble of extracted outliers (nof1disease) from 4.0json
```{r input up gene data sample 5}

# 1. Repeat step #1 for sample 5, changing TH_id
nof1_up_data_5 <- read_tsv("Sample 1-6 nof1 Outlier Results.tsv", show_col_types = FALSE) %>%
  filter(direction == "up") %>%
  filter(TH_id == "TH50_5282_S01") %>% 
  filter(!grepl("^RP", gene))
  
# Add expression levels from outlier_results$sample (lets call, data set B)
# 2. Repeat step 2, changing TH_id
json_file4.0_5 <- file.path(getwd(), "TH50_5282_S01.4.0.json")

json_data4.0_5 <- fromJSON(json_file4.0_5)

up_outlier_expression_df_5 <- tibble(gene=names(json_data4.0_5$outlier_results$sample), expression=unlist(json_data4.0_5$outlier_results$sample))

print(up_outlier_expression_df_5)

# 3. Merge the two datasets (inner_join for overlapping genes) (A+B)
up_merged_data_5 <- inner_join(nof1_up_data_5, up_outlier_expression_df_5, by = "gene")

# 4. Select columns 
up_final_data_5 <- up_merged_data_5 %>%
  select(gene, direction, expression)

print(up_final_data_5)

# 5. Merge the threshold data with the expression data
up_final_merged_data_5 <- inner_join(up_final_data_5, threshold_up_outlier_genes_df, by = "gene") %>%
  select(gene, expression, threshold, direction)

# Style kable table 
kable(up_final_merged_data_5) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "500px")

# Export as TSV
write.table(up_final_merged_data_5, "Sample-5-Up-Outliers-RP-Filtered.tsv", sep = "\t", row.names = FALSE, quote = FALSE)

```

## Create Up Outlier Gene tibble for Sample 6, including gene and expression level (4.0json), and merge with tibble of extracted outliers (nof1disease) from 4.0json
```{r input up gene data sample 6}

# 1. Repeat step #1 for sample 6, changing TH_id
nof1_up_data_6 <- read_tsv("Sample 1-6 nof1 Outlier Results.tsv", show_col_types = FALSE) %>%
  filter(direction == "up") %>%
  filter(TH_id == "TH50_5283_S01") %>% 
  filter(!grepl("^RP", gene))
  
# Add expression levels from outlier_results$sample (lets call, data set B)
# 2. Repeat step 2, changing TH_id
json_file4.0_6 <- file.path(getwd(), "TH50_5283_S01.4.0.json")

json_data4.0_6 <- fromJSON(json_file4.0_6)

up_outlier_expression_df_6 <- tibble(gene=names(json_data4.0_6$outlier_results$sample), expression=unlist(json_data4.0_6$outlier_results$sample))

print(up_outlier_expression_df_6)

# 3. Merge the two datasets (inner_join for overlapping genes) (A+B)
up_merged_data_6 <- inner_join(nof1_up_data_6, up_outlier_expression_df_6, by = "gene")

# 4. Select columns 
up_final_data_6 <- up_merged_data_6 %>%
  select(gene, direction, expression)

print(up_final_data_6)

# 5. Merge the threshold data with the expression data
up_final_merged_data_6 <- inner_join(up_final_data_6, threshold_up_outlier_genes_df, by = "gene") %>%
  select(gene, expression, threshold, direction)

# Style kable table 
kable(up_final_merged_data_6) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "500px")

# Export as TSV
write.table(up_final_merged_data_6, "Sample-6-Up-Outliers-RP-Filtered.tsv", sep = "\t", row.names = FALSE, quote = FALSE)

```

## Samples 1-6 COMPLETE UP Outliers 
```{r add count}

# Add TH_id and Fay_id colums to each of 6 sample df's to keep track when merging
up_final_merged_data_1 <- up_final_merged_data_1 %>% mutate(TH_id = "TH50_5178_S01", Fay_id = "Sample 1")
up_final_merged_data_2 <- up_final_merged_data_2 %>% mutate(TH_id = "TH50_5179_S01", Fay_id = "Sample 2")
up_final_merged_data_3 <- up_final_merged_data_3 %>% mutate(TH_id = "TH50_5280_S01", Fay_id = "Sample 3")
up_final_merged_data_4 <- up_final_merged_data_4 %>% mutate(TH_id = "TH50_5281_S01", Fay_id = "Sample 4")
up_final_merged_data_5 <- up_final_merged_data_5 %>% mutate(TH_id = "TH50_5282_S01", Fay_id = "Sample 5")
up_final_merged_data_6 <- up_final_merged_data_6 %>% mutate(TH_id = "TH50_5283_S01", Fay_id = "Sample 6")

# Combine all 6 samples expression data
all_up_final_merged_data <- bind_rows(
  up_final_merged_data_1,
  up_final_merged_data_2,
  up_final_merged_data_3,
  up_final_merged_data_4,
  up_final_merged_data_5,
  up_final_merged_data_6
  )

# Calculate gene count/presence across sample cohort 
up_gene_counts <- all_up_final_merged_data %>%
  group_by(gene) %>%
  summarize(count = n(), .groups = "drop")

# Add counts to combined data
final_up_data_with_counts <- left_join(all_up_final_merged_data, gene_counts, by = "gene")

# Style kable table 
kable(final_up_data_with_counts) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "500px")

# Export as TSV
write.table(final_up_data_with_counts, "Samples-1-to-6-Up-Outliers.tsv", sep = "\t", row.names = FALSE, quote = FALSE)
```

