---
title: "ReactomePA Sample 1-6"
author: "Kaylee Delgado"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  word_document: default
---

Description from Bioconductor: "This package provides functions for pathway analysis based on REACTOME pathway database. It implements enrichment analysis, gene set enrichment analysis and several functions for visualization.
This package is not affiliated with the Reactome team."

Explainer File : https://bioconductor.org/packages/release/bioc/manuals/ReactomePA/man/ReactomePA.pdf

## Install Packages
```{r packages, results='hide', message=FALSE}

#install packages
CRANpkg <- function (pkg) {
    cran <- "https://CRAN.R-project.org/package"
    fmt <- "[%s](%s=%s)"
    sprintf(fmt, pkg, cran, pkg)
}

library(dplyr)
library(org.Hs.eg.db)
library(DOSE)
library(ReactomePA)
library(AnnotationDbi)
library(BiocGenerics)
library(stats4)
library(knitr)
library(conflicted)
library(readr)
library(formatR)
library(stringr)
library(kableExtra)
library(evaluate)
library(rmarkdown)
```

## File Input
```{r}

# Read the TSV file into a data frame
outlier_data <- read.table("Sample 1-6 nof1 Outlier Results.tsv", sep = "\t", header = TRUE)
```

## Pathways Enrichment:
### Down Outliers for TH50_5178_S01
```{r, message=FALSE}
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)

# Input down outlier genes
down_outliers_78 <- outlier_data %>%
  filter(TH_id == "TH50_5178_S01", direction == "down") %>%
  select(gene)

# First turn outlier gene name list into gene ID list
down_gene_IDs_78 <- mapIds(org.Hs.eg.db, keys = down_gene_names_78, column = "ENTREZID", keytype = "SYMBOL")

# Second perform Reactome Pathway Enrichment
down_enrich_pathway_78 <- enrichPathway(down_gene_IDs_78, organism = "human", pvalueCutoff = 0.05, readable = TRUE)

# Read results as a dataframe
down_ep_78_results <- as.data.frame(down_enrich_pathway_78)

kable(down_ep_78_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Up Outliers for TH50_5178_S01
```{r, message=FALSE}

#repeat above steps for up outliers
up_outliers_78 <- outlier_data %>%
  filter(TH_id == "TH50_5178_S01", direction == "up") %>%
  select(gene)

up_gene_names_78 <- up_outliers_78$gene

up_gene_IDs_78 <- mapIds(org.Hs.eg.db, keys = up_gene_names_78, column = "ENTREZID", keytype = "SYMBOL")

up_enrich_pathway_78 <- enrichPathway(up_gene_IDs_78, organism = "human", pvalueCutoff = 0.05, readable = TRUE)

up_ep_78_results <- as.data.frame(up_enrich_pathway_78)

kable(up_ep_78_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Down Outliers for TH50_5179_S01
```{r, message=FALSE}

#now repeat for all samples
down_outliers_79 <- outlier_data %>%
  filter(TH_id == "TH50_5179_S01", direction == "down") %>%
  select(gene)

down_gene_IDs_79 <- mapIds(org.Hs.eg.db, keys = down_gene_names_79, column = "ENTREZID", keytype = "SYMBOL")

down_enrich_pathway_79 <- enrichPathway(down_gene_IDs_79, organism = "human", pvalueCutoff = 0.05, readable = TRUE)

down_ep_79_results <- as.data.frame(down_enrich_pathway_79)

kable(down_ep_79_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Up Outliers for TH50_5179_S01
```{r, message=FALSE}
up_outliers_79 <- outlier_data %>%
  filter(TH_id == "TH50_5179_S01", direction == "up") %>%
  select(gene)

up_gene_IDs_79 <- mapIds(org.Hs.eg.db, keys = up_gene_names_79, column = "ENTREZID", keytype = "SYMBOL")

up_enrich_pathway_79 <- enrichPathway(up_gene_IDs_79, organism = "human", pvalueCutoff = 0.05, readable = TRUE)

up_ep_79_results <- as.data.frame(up_enrich_pathway_79)

kable(up_ep_79_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Down Outliers for TH50_5280_S01
```{r, message=FALSE}
down_outliers_80 <- outlier_data %>%
  filter(TH_id == "TH50_5280_S01", direction == "down") %>%
  select(gene)

down_gene_IDs_80 <- mapIds(org.Hs.eg.db, keys = down_gene_names_80, column = "ENTREZID", keytype = "SYMBOL")

down_enrich_pathway_80 <- enrichPathway(down_gene_IDs_80, organism = "human", pvalueCutoff = 0.05, readable = TRUE)

down_ep_80_results <- as.data.frame(down_enrich_pathway_80)

kable(down_ep_80_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Up Outliers for TH50_5280_S01
```{r, message=FALSE}
up_outliers_80 <- outlier_data %>%
  filter(TH_id == "TH50_5280_S01", direction == "up") %>%
  select(gene)

up_gene_IDs_80 <- mapIds(org.Hs.eg.db, keys = up_gene_names_80, column = "ENTREZID", keytype = "SYMBOL")

up_enrich_pathway_80 <- enrichPathway(up_gene_IDs_80, organism = "human", pvalueCutoff = 0.05, readable = TRUE)

up_ep_80_results <- as.data.frame(up_enrich_pathway_80)

kable(up_ep_80_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Down Outliers for TH50_5281_S01
```{r, message=FALSE}
down_outliers_81 <- outlier_data %>%
  filter(TH_id == "TH50_5281_S01", direction == "down") %>%
  select(gene)

down_gene_IDs_81 <- mapIds(org.Hs.eg.db, keys = down_gene_names_81, column = "ENTREZID", keytype = "SYMBOL")

down_enrich_pathway_81 <- enrichPathway(
  down_gene_IDs_81,
  organism = "human",
  pvalueCutoff = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.2,
  minGSSize = 10,
  maxGSSize = 500,
  readable = TRUE
)

down_ep_81_results <- as.data.frame(down_enrich_pathway_81)

kable(down_ep_81_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Up Outliers for TH50_5281_S01
```{r, message=FALSE}
up_outliers_81 <- outlier_data %>%
  filter(TH_id == "TH50_5281_S01", direction == "up") %>%
  select(gene)

up_gene_IDs_81 <- mapIds(org.Hs.eg.db, keys = up_gene_names_81, column = "ENTREZID", keytype = "SYMBOL")

up_enrich_pathway_81 <- enrichPathway(
  up_gene_IDs_81,
  organism = "human",
  pvalueCutoff = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.2,
  minGSSize = 10,
  maxGSSize = 500,
  readable = TRUE
)

up_ep_81_results <- as.data.frame(up_enrich_pathway_81)
```

### Down Outliers for TH50_5282_S01
```{r, message=FALSE}
down_outliers_82 <- outlier_data %>%
  filter(TH_id == "TH50_5282_S01", direction == "down") %>%
  select(gene)

down_gene_IDs_82 <- mapIds(org.Hs.eg.db, keys = down_gene_names_82, column = "ENTREZID", keytype = "SYMBOL")

down_enrich_pathway_82 <- enrichPathway(
  down_gene_IDs_82,
  organism = "human",
  pvalueCutoff = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.2,
  minGSSize = 10,
  maxGSSize = 500,
  readable = TRUE
)

down_ep_82_results <- as.data.frame(down_enrich_pathway_82)

kable(down_ep_82_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Up Outliers for TH50_5282_S01
```{r, message=FALSE}
up_outliers_82 <- outlier_data %>%
  filter(TH_id == "TH50_5282_S01", direction == "up") %>%
  select(gene)

up_gene_IDs_82 <- mapIds(org.Hs.eg.db, keys = up_gene_names_82, column = "ENTREZID", keytype = "SYMBOL")

up_enrich_pathway_82 <- enrichPathway(
  up_gene_IDs_82,
  organism = "human",
  pvalueCutoff = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.2,
  minGSSize = 10,
  maxGSSize = 500,
  readable = TRUE
)

up_ep_82_results <- as.data.frame(up_enrich_pathway_82)

kable(up_ep_82_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Down Outliers for TH50_5283_S01
```{r, message=FALSE}
down_outliers_83 <- outlier_data %>%
  filter(TH_id == "TH50_5283_S01", direction == "down") %>%
  select(gene)

down_gene_IDs_83 <- mapIds(org.Hs.eg.db, keys = down_gene_names_83, column = "ENTREZID", keytype = "SYMBOL")

down_enrich_pathway_83 <- enrichPathway(
  down_gene_IDs_83,
  organism = "human",
  pvalueCutoff = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.2,
  minGSSize = 10,
  maxGSSize = 500,
  readable = TRUE
)

down_ep_83_results <- as.data.frame(down_enrich_pathway_83)

kable(down_ep_83_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```

### Up Outliers for TH50_5283_S01
```{r, message=FALSE}
up_outliers_83 <- outlier_data %>%
  filter(TH_id == "TH50_5283_S01", direction == "up") %>%
  select(gene)

up_gene_IDs_83 <- mapIds(org.Hs.eg.db, keys = up_gene_names_83, column = "ENTREZID", keytype = "SYMBOL")

up_enrich_pathway_83 <- enrichPathway(
  up_gene_IDs_83,
  organism = "human",
  pvalueCutoff = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.2,
  minGSSize = 10,
  maxGSSize = 500,
  readable = TRUE
)

up_ep_83_results <- as.data.frame(up_enrich_pathway_83)

kable(up_ep_83_results) %>%
  kable_styling(full_width = FALSE, position = "left")
```